**#Prediction Assignment Writeup"**


The goal of this project to predict the the manner of people did during exercise using data from http://groupware.les.inf.puc-rio.br/har. 

##Preparing the Data (Data Cleaning)
###Training Dataset
The training dataset is from https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv.
There are 19622 rows of training datasets, which means there are 19622 dataset for training, with 160 variables (or features).

###Testing dataset
The testing dataset is from https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv
There are 20 rows of datasets, which means there are 20 dataset for training, with 160 variables (or features).


###Strategy
This project will partition the training dataset into 70% for training, and 30% for validation on creating the machine learning model. 


Not all 160 features are needed, so we need to clean the dataset by removing some columns. For example, columns with NA values, DIV/0, empty columns, and column with people names that is not correlated with the expected predicted classes.

```{r}
## Data loading
setwd("..")
pmlTrain<-read.csv("pml-training.csv", header=T, na.strings=c("NA", "#DIV/0!"))
pmlTest<-read.csv("pml-testing.csv", header=T, na.string=c("NA", "#DIV/0!"))

## NA exclusion for all available variables
noNApmlTrain<-pmlTrain[, apply(pmlTrain, 2, function(x) !any(is.na(x)))] 
dim(noNApmlTrain)

## variables with user information, time and undefined
cleanpmlTrain<-noNApmlTrain[,-c(1:8)]
dim(cleanpmlTrain)

## 20 test cases provided clean info - Validation data set
cleanpmltest<-pmlTest[,names(cleanpmlTrain[,-52])]
dim(cleanpmltest)
```

 The training dataset into 70% for training, and 30% for validation on creating the machine learning model, using these codes:
 
```{r}
#Create Partition
library(caret)
inTrain<-createDataPartition(y=cleanpmlTrain$classe, p=0.70,list=F)
training<-cleanpmlTrain[inTrain,] 
validate<-cleanpmlTrain[-inTrain,] 

#Training and validate set dimensions
dim(training)
dim(validate)

```

##Training
The training dataset were trained using Random Forest Trees. 3 fold cross validation was used control the model. 
Random forest trees were generated for the training dataset using cross-validation. Then the generated algorithm was examined under the partitioned training set to examine the accuracy and estimated error of prediction. 

By using 51 features for five classes using cross-validation at a 3-fold an accuracy of 99.54% with a 95% CI [0.9933, 0.997] was achieved accompanied by a Kappa value of 0.9942.

```{r}
library(caret)
set.seed(786541)
pmlTrainControl<-trainControl(method="cv", number=3, allowParallel=T, verbose=T)
pmlTrainedModel<-train(classe~.,data=training, method="rf", trControl=pmlTrainControl, verbose=F)
```


##Validating
The  trained model is then tested on the validation dataset (the 30% data partitioned from original training dataset.
```{r}
predrf<-predict(pmlTrainedModel, newdata=validate)
confusionMatrix(predrf, validate$classe)

```


#Results and Conclusion
The trained model is then used to make prediction on the 20 dataset given at the beginning of the project.

```{r}
pred20<-predict(pmlTrainedModel, newdata=cleanpmltest)

```

